---
title: "deduplicate-complaints-dk"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review Duplicate Complaints

For complaints with duplicate IDs or notes indicating they are a duplicate, seeing if they refer to the same property/complaint. 

Original project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212241865138317?focus=true),
though note that scope has changed somewhat so information is outdated.

Result of this script should be:

1. A file with de-duplicated complaints when their IDs refer to the same complaint
2. A file with remaining duplicate IDs that seem to refer to different complaints,
which may need review


TO DO:

* Combine information across duplicates
* Check that when one of name/address matches and the other doesnt that they are actually the same place

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
source("0-fns-dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260623" # date of file exports

set.seed(111)
```

### Load Data
```{r}
comps <- readRDS(paste(data_dir,"complaints_duplicate_id_dk.RDS",sep="/"))

comps <- comps |>
  group_by(complaint_id) |>
  arrange(complaint_id) |>
  mutate(comp_num=row_number())

starting_comps <- nrow(comps)
```


## Flag true duplicates

### Date of complaint
NOTE: This should be used to indicate whether a complaint is a duplicate or not, but will not be a criteria for
whether or not to use for ID matching. Even if complaints have different dates, ID matching is a valid approach
if they refer to the same location.


```{r}
day_mark <- 2
comps <- comps |>
  mutate(
    same_date = date_complaint_received ==
      lag(date_complaint_received, 1, order_by = comp_num),
    close_date = as.numeric(
      difftime(date_complaint_received,
        lag(date_complaint_received, 1, order_by = comp_num),
        units = "days"
      )
    ) <= day_mark
  )

```

### Location of complaint
```{r}
comps <- comps |> 
  mutate(
    last_propname=lag(property_name, 1, order_by = comp_num),
    last_addr=lag(property_address, 1, order_by = comp_num),    
  ) |>
  mutate(
    same_prop=property_name==last_propname,
    same_addr=property_address==last_addr,
    same_loc=case_when(
      comp_num==1 ~ NA,
      is.na(same_prop) & is.na(same_addr) ~ FALSE,
      same_prop | same_addr ~ TRUE,
      TRUE ~ FALSE
    )
  )
    
```

Space for review
```{r}
#comps |>
#  select(original_id:complaint_id,property_name,property_address,property_address_orig,same_prop:same_loc) |>
#  View()
```



Space for ad hoc reviews
```{r}
#comps |>
#  filter(complaint_id=="1010640") |>  
  #filter(complaint_id=="1747318") |>
  #filter(complaint_id=="1063473") |>
#  select(property_name,property_address,last_propname,last_addr,same_prop,same_addr,same_loc)
```


### Complainant name
Potential future addition

### Similarity in notes
Potential future addition


### Flag duplicates
Not needed when just identifying different properties but will be needed for other work on duplicates 
```{r}
#comps <- comps |>
#  mutate(same_as_above=same_date & same_loc) # could also use close_date
```



## Tag duplicates

### Separate same vs different location

Problematic duplicates (e.g. those that seem to refer to different properties)
from those that can be safely de-duplicated
```{r}
comps <- comps |>
  mutate(has_problematic_dupes=!as.logical(min(same_loc,na.rm = TRUE)))


comps_diff_loc <- comps |>
  filter(has_problematic_dupes) |>
  select(1:diff_source)  

comps_same_loc <- comps |>
  filter(!has_problematic_dupes) |>
  select(1:diff_source)  


```

Space for review
```{r}
#comps_diff_loc |>
#  select(complaint_id,property_name,property_address,property_address_orig,same_prop:last_col()) |>
#  View()
```



### TO DO: Compare and combine information
Removing the below for now as we should not fully deduplicate
```{r}
#comps_deduped <- comps |>
#  distinct(complaint_id,
#           date_complaint_received,
#           property_name,
#           .keep_all = TRUE) |>
#  distinct(complaint_id,
#           date_complaint_received,
#           property_address,
#           .keep_all = TRUE)
```

```{r}
nrow(comps) 
nrow(comps_diff_loc)
nrow(comps_same_loc)
```


## Export
```{r}
# From old code, unused files
#saveRDS(comps_with_dupes,paste(data_dir,"duplicate_complaints_review.xlsx",sep="/"))
#saveRDS(comps_deduped,paste(data_dir,"complaints_deduplicated_dk.RDS",sep="/"))


saveRDS(comps_diff_loc,paste(data_dir,"duplicate_complaints_diffloc.RDS",sep="/"))
saveRDS(comps_same_loc,paste(data_dir,"duplicate_complaints_sameloc.RDS",sep="/"))

```

