---
title: "connect-properties-311"
author: "Dylan Knaggs"
date: "2025-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Properties with Complaints by ID

Combine Complaints with the properties dataset [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212050138239287?focus=true)


NOTE: Something that needs to be accounted for within all matches is the fact that
many properties have a duplicate, where one record is an inactive property and the other
is an active property, i.e. the property was used as a rental, stopped being so 
(perhaps because of being sold), then was registered again. Currently there's 
no accounting for when the property was registered vs when the complaint occurred

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(tidygeocoder)
library(readxl)
library(writexl)
library(janitor)
library(here)
source("0-fns-dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260623" # date of file exports

```


## Load Data

### Properties
```{r}
props <- readRDS(paste(data_dir,"properties_parcel_match_dk.RDS",sep="/")) |>
  mutate(props=1)
```

### Parcels
```{r}
parcels <- read_csv(paste0(data_dir,"/parcels_",date_str,".csv")) |>
  clean_names() |>
  mutate(parcels=1)

```


### 311 Calls
```{r}
calls_new <- read_csv(
    paste0(data_dir,"/311_calls_post_mar21_",date_str,".csv")
  ) |>
  clean_names() |>
  mutate(calls=1)

calls_old <- read_csv(paste(data_dir,"311_calls_pre_mar21_20251210.csv",
                   sep="/")) |>
  clean_names() |>
  filter(work_group=="Health-Healthy Homes-") |>
  mutate(calls=1)

```

### Access Complaints
```{r}
comps <- bind_rows(
  readRDS(paste(data_dir,"complaints_unique_id_dk.RDS",sep="/")),
  readRDS(paste(data_dir,"duplicate_complaints_sameloc.RDS",sep="/"))
) |>
  mutate(comps=1)
```



## Align Columns

### 311 Calls

Column types
```{r}
calls_new <- calls_new |>
  mutate(
    id=as.character(id),
    pin=as.character(pin),
    open_date=as.Date(open_date,"%m/%d/%Y"),    
    resolved_date=as.Date(resolved_date,"%m/%d/%Y"),    
  )

calls_old <- calls_old |>
  mutate(
    case_id=as.character(case_id),
    parcel_id_no=as.character(parcel_id_no),
    creation_date=as.Date(creation_date,"%m/%d/%Y"),    
    closed_date=as.Date(closed_date,"%m/%d/%Y"),    
  )

```

Column names (and bind new+old calls)
```{r}
calls <- bind_rows(
  calls_new |> 
    select(id,pin,status,open_date,resolved_date,
           address,latitude,longitude,open_date,calls)|>
    rename(property_address=address,complaint_id=id),
  
  calls_old |> 
    select(case_id,parcel_id_no,status,creation_date,closed_date,
           street_address,latitude,longitude,calls) |>
    rename(complaint_id=case_id,pin=parcel_id_no,open_date=creation_date,
           resolved_date=closed_date,property_address=street_address)
  )
  
```


### parcels
```{r}
parcels <- parcels |>
  select(kivapin,apn,address,landusecode,parcels) |>
  rename(pin=kivapin,parcel_num=apn,property_address=address)

```

```{r}
parcels <- parcels |>
  mutate(pin=as.character(pin))

```


## Check Uniqueness 

### Properties
```{r}
props |>
  mutate(parcel_ct=n(),.by=parcel_num) |>
  filter(parcel_ct>1) |>
  arrange(parcel_num)
```

NOTE:
there are a small number of properties with duplicate parcel numbers. These
should be handled in deduplicate-properties.Rmd, but going to remove here for now,
keeping the most recent.

```{r}
props <- props |>
  arrange(desc(application_date)) |>
  distinct(parcel_num,.keep_all = TRUE)
```


### Parcels
```{r}
parcels |> 
  count_dat(pin) |>
  filter(n>1)

parcels |> 
  count_dat(parcel_num) |>
  filter(n>1)
```


There are also a small number of parcels with duplicate parcel numbers or 
pins. Deduplicating arbitrarily for now.

```{r}
parcels <- parcels |> 
  distinct(parcel_num,.keep_all = TRUE) |>
  distinct(pin,.keep_all = TRUE)  
```


### 311 Calls
```{r}
calls |> 
  count_dat(complaint_id) |>
  filter(n>1)
```


### Access Complaints

```{r}
comps |> 
  count_dat(complaint_id) |>
  filter(n>1)
```


## Join datasets

### Access complaints to calls
Note that this uses the "root id" which sets "xxxx.1" to "xxxx" and is how the
id would be stored in 311
```{r}
comps_311 <- comps |>
  left_join(calls, by=c("root_id"="complaint_id"),suffix=c("","_311"),  
            relationship="many-to-one") 

unmatched_to_311 <- comps_311 |>
  filter(is.na(calls))

comps_311 <- comps_311 |>
  filter(!is.na(calls))

```

### Include parcels
```{r}
comps_par <- comps_311 |>
  left_join(parcels,by="pin",suffix=c("","_par"),
            relationship = "many-to-one")

unmatched_to_par <- comps_par |>
  filter(is.na(parcels))

comps_par <- comps_par |>
  filter(!is.na(parcels))

```

### Include properties
```{r}
comps_prop <- comps_par |>
  left_join(props,by="parcel_num",suffix=c("","_prop"),
            relationship = "many-to-one")

unmatched_to_prop <- comps_prop |>
  filter(is.na(props))

comps_prop <- comps_prop |>
  filter(!is.na(props))


```


## Review
Confirm row counts and check number of matches by stage 
```{r}
all_unmatched <- bind_rows(
  unmatched_to_311,
  unmatched_to_par,
  unmatched_to_prop
)

# Number matched and unmatched
paste("Total Matched:",nrow(comps_prop))
paste("Total Unmatched:",nrow(all_unmatched))
print("")

# Unmatched at each stage
paste("Unmatched from Complaint to 311:",nrow(unmatched_to_311))
paste("Unmatched from 311 to Parcels:",nrow(unmatched_to_par))
paste("Unmatched from Parcels to HHRIP Properties:",
      sum(all_unmatched$parcels,na.rm=TRUE))


# confirm all properties are accounted for
print("")
paste("Unmatched+Matched:",nrow(all_unmatched)+nrow(comps_prop))
paste("Starting # of Complaints:",nrow(comps))

```

### Review nonmatches
Biggest issues occurs from parcel to property match. Seeing if there are trends in mismatches over time
```{r}
unmatched_to_prop |> 
  ungroup() |>
  mutate(month=floor_date(date_complaint_received,unit="month")) |>
  dplyr::count(month) |>
  plot(type="b")

```


```{r}
all_unmatched |>
  filter(parcels==1) |>
  select(complaint_id,property_address,property_name,date_complaint_received,
         parcel_num,pin,comps,calls,parcels) |>
  arrange(desc(date_complaint_received))


```



Based on spot checks against properties and the parcel viewer, reasons why non-matches could occur here:
1. duplicate parcel, though in many cases one of these is active and the other is inactive
  Could do an inequality join based on complaint/registration date (though not sure
  complaint date will always be after registration date)
2. Missing parcel numbers
3. Might not be registered? (608 E 43rd)
4. Different addresses/parcels (RIDGEVIEW HEIGHTS APARTMENTS)


### Review
```{r}
comps_prop |>
  select(complaint_id,property_address,property_address_prop,
         property_address_311,property_address_par,property_name,property_name_prop,
         date_complaint_received,parcel_num,pin) |>
#  filter(is.na(property_name) & property_address!=property_address_prop) |> #testing some different conditions
  arrange(desc(date_complaint_received)) |>
  View()


```



## Export
### Nonmatches
```{r}
all_unmatched |> 
  saveRDS(paste(data_dir,"comps_unmatched_by_id.RDS",sep="/"))
```


### Matches
```{r}
comps_prop |>
  relocate(property_id,owner_id,property_name_prop,.after=complaint_id) |>
  saveRDS(paste(data_dir,"comps_matched_by_id.RDS",sep="/"))

comps_prop |>
  ungroup () |>
  relocate(property_id,owner_id,property_name_prop,.after=complaint_id) |>
  select(any_of(keep_cols_match)) |>
  write_xlsx(paste(outputs_dir,"matched_comps_id.xlsx",sep="/"))
  
```

