---
title: "connect-properties-complaints-location"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Complaints to Properties by Location

Match complaints to properties based on property name and address. This considers the 
complaints that were not previously matched by ID.

Inputs:
1. Complaints
  A. Unmatched by ID
  B. Duplicated by ID (with two different locations)
  C. Missing ID
2. Cleaned Property file

Result of this script should be:

1. Complaints matched to properties by address/property name
2. Unmatched Complaints

## INIT
```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
library(reclin2)
library(wordcloud)
library(sf)
source("0-fns-dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260623" # date of file exports

```

## Read in data

### Complaints
```{r}
comps_unmatched_orig <- readRDS(paste(
  data_dir,"comps_unmatched_by_id.RDS",sep="/")) |>
  mutate(from="UNMATCHED",source="COMPS")

comps_miss_id_orig <- readRDS(paste(
  data_dir,"complaints_missing_id_dk.RDS",sep="/")) |>
  mutate(from="MISSING",source="COMPS")


comps_dupe_id_orig <- readRDS(paste(
  data_dir,"duplicate_complaints_diffloc.RDS",sep="/")) |>
  mutate(from="DUPE",source="COMPS")

# combine
comps <- bind_common_cols(
  comps_dupe_id_orig,comps_miss_id_orig,comps_unmatched_orig
)

comps_start <- comps
```


Add an ID for tracking purposes
```{r}
comps$match_id <- 1:nrow(comps)

```



### Properties
```{r}
props <- read_all_props()

```


### String Cleaning 
```{r}
comps <- comps |> clean_prop_names()
props <- props |> clean_prop_names()
  
```

### Property duplicates
There are still some duplicate properties with the same application date. 
Deduplicating here but should flag to HHRIP/move deduplication to 
property cleaning script
```{r}
props <- props |>
  arrange(from) |>
  distinct(property_name,application_date,.keep_all = TRUE)

```


## Exact Matches

"Exact Matches" isn't necessarily the best name because it removes some common 
words that are inconsistently put in property names ("THE", "APARTMENTS", "TOWNHOMES", etc.).

### Property Names

Get matches
```{r}
comps_name_match <- comps |> 
  left_join(props,
            by=join_by(
              property_name==property_name,
#              closest(date_complaint_received>application_date)
            ),
            na_matches = "never",
            keep=TRUE,
            suffix=c("_comp","_prop"),
            relationship = "many-to-many"
  ) |>
  filter(!is.na(property_id))


```

There are some cases where there are multiple matches because of the same
property having an active and inactive version. Keeping the record where
the complaint date is closest to the registration date (regardless of whether its before or after,
because all properties can have complaints). Should review this logic with HHRIP.

```{r}
comps_name_match <- comps_name_match |>
  mutate(comp_app_diff=abs(date_complaint_received-application_date)) |>
  arrange(match_id,comp_app_diff) |>
  distinct(match_id,.keep_all = TRUE) |>
  select(-comp_app_diff)

print(paste("Number of complaints matched by prop_name:",nrow(comps_name_match)))

```


Separate nonmatches
```{r}
comps <- comps |>
  anti_join(comps_name_match,by="match_id")
print(paste("Number of complaints still unmatched:",nrow(comps)))
```

### Addresses
```{r}
comps_addr_match <- comps |>
  left_join(props, 
            by=join_by(
              property_address==property_address,
#              closest(date_complaint_received>application_date) 
            ),
            na_matches = "never",
            keep=TRUE,
            suffix=c("_comp","_prop"),
            relationship = "many-to-many"
) |>
  filter(!is.na(property_id))
```


There are some cases where there are multiple matches because of the same
property having an active and inactive version. Keeping the record where
the complaint date is closest to the registration date (regardless of whether its before or after,
because all properties can have complaints). Should review this logic with HHRIP.

```{r}
comps_addr_match <- comps_addr_match |>
  mutate(comp_app_diff=abs(date_complaint_received-application_date)) |>
  arrange(match_id,comp_app_diff) |>
  distinct(match_id,.keep_all = TRUE) |>
  select(-comp_app_diff)

print(paste("Number of complaints matched by address:",nrow(comps_addr_match)))

```


Separate nonmatches
```{r}
comps <- comps |>
  anti_join(comps_addr_match,by="match_id")

print(paste("Number of complaints still unmatched:",nrow(comps)))

```

## Counts
```{r}
nrow(comps) # still unmatched
nrow(comps_name_match) # matched by name
nrow(comps_addr_match) # matched by address
nrow(comps)+nrow(comps_name_match)+nrow(comps_addr_match) # total of all cats
nrow(comps_start) # total at start (should be the same as above)


```

```{r}
length(unique(comps$match_id))==length(comps$match_id)
length(unique(comps_addr_match$match_id))==length(comps_addr_match$match_id)
length(unique(comps_name_match$match_id))==length(comps_name_match$match_id)
```

```{r}
comps_name_match |>
  ungroup() |>
  mutate(n=n(),.by=match_id) |>
  filter(n>1) |>
  View()
#  distinct(property_name)
```


## Nonmatches

Reviewing nonmatches
```{r}
comps |>
  select(match_id,complaint_id,date_complaint_received,inspection_date,
         hud_complaint,
         complaint_outcome_comments,complainant_name,property_name,
         property_address,geo_address,property_address_orig,property_zip_code,
         arcgis_location.x,arcgis_location.y) |> 
  filter(date_complaint_received>=as.Date("2025-06-01")) |>
  View()
```

ad hoc reviews
```{r}
props |>
  filter(str_detect(property_address,"DENVER")) |> arrange(property_address)
#  filter(str_detect(property_name,"FREN"))


#12509 E 53RD TER, KANSAS CITY, MISSOURI, 64133 | exists as 12509 E 53rd St
#7351 NW OLD TIFFANY SPRINGS RD, BUILDING 3, KANSAS CITY, MISSOURI, 64153 | nearby (7441 and others) but not exact match
# 4840 PASEO BLVD, KANSAS CITY, MISSOURI, 64110 | looks like part of UNIVERSITY MEADOWS
# 3200 DENVER AVE, KANSAS CITY, MISSOURI, 64128 | looks like part of RIDGEVIEW HEIGHTS

```


Nonmatches over time
```{r}
comps |>
  ggplot(mapping=aes(x=rollback(date_complaint_received))) +
  geom_bar()

```


Some notes on other things to check
Address ranges
fuzzy matching for property names
management office is the property address (probably for comps)
  needs hhrips review
  do they have a list of mgmt companies and addresses?
more?


## Export
Review files
```{r}
saveRDS(comps,
        paste0(data_dir,"/unmatched_comps_id_propname_address.RDS"))

write_xlsx(comps |> select(any_of(keep_cols_nonmatch)),
           paste0(outputs_dir,"/unmatched_comps_id_propname_address.xlsx"))

write_xlsx(comps_name_match |> select(any_of(keep_cols_match)),
           paste0(outputs_dir,"/matched_comps_propname.xlsx"))

write_xlsx(comps_addr_match |> select(any_of(keep_cols_match)),
           paste0(outputs_dir,"/matched_comps_address.xlsx"))
```


