---
title: "connect-properties-complaints-location"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Complaints to Properties by Location

Match complaints to properties based on factors with less precision than the previous stages. Potential matches here need manual confirmation.

Inputs:
1. Complaints still unmatched after previous stages
2. All properties

Result of this script should be:
1. Complaints with best-guess matches to properties ready for manual review
2. Unmatched Complaints

## INIT
```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
library(reclin2)
library(wordcloud)
library(sf)
source("0-fns-dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260623" # date of file exports

```

## Read in data
```{r}
comps <- readRDS(paste0(data_dir,"/unmatched_comps_id_propname_address.RDS"))
props <- read_all_props()
```




## Nearby complaint-address pairs
Not sure how useful/valid this will end up being

Convert to sf
```{r}
comps_sf <- comps |> 
  filter(!is.na(arcgis_location.x) & !is.na(property_address)) |>
  st_as_sf(coords=c("arcgis_location.x","arcgis_location.y")) |>
  select(complaint_id,property_name,property_address)

st_crs(comps_sf) <- 4326

props_sf <- props |>
  filter(!is.na(arcgis_location.x) & !is.na(property_address)) |>
  st_as_sf(coords=c("arcgis_location.x","arcgis_location.y")) |>
  select(property_id,property_name,property_address) |>
  rename_with(~ paste0(.x,"_prop"))
st_crs(props_sf) <- 4326

```

Spatial join on distance (match properties within 50m)
```{r}
comps_near_sf <- st_join(comps_sf,props_sf,
                         join=st_is_within_distance,dist=50,
                         left=FALSE) |> 
  select(contains("_id"),contains("address"),contains("name"))
  
```

### Review 
Unique matches
```{r}
dist_match_unique <- comps_near_sf |>
  mutate(n_matches=n(),.by=complaint_id) |>
  filter(n_matches==1) 

dist_match_unique |> nrow()

#dist_match_unique |> View()
```

Non-unique matches
```{r}
mult <- comps_near_sf |>
  mutate(n_matches=n(),.by=complaint_id) |>
  filter(n_matches>1) |>
  arrange(complaint_id, property_address, property_address_prop) 
  
mult |> nrow()  
#mult |>  View()
```

Non-matches plus non-unique matches
```{r}
inconc_add <- comps |>
  filter(!complaint_id %in% dist_match_unique$complaint_id) |>
#  filter(!complaint_id %in% comps_near_sf$complaint_id | 
#           complaint_id %in% mult$complaint_id) |>
  select(complaint_id,date_complaint_received,inspection_date,hud_complaint,
         complaint_outcome_comments,complainant_name,property_name,
         property_address,geo_address,property_address_orig,property_zip_code,
         arcgis_location.x,arcgis_location.y)  

nrow(inconc_add)
```

## Fuzzy Matching on property name

Still needs development

```{r}
comps_w_propnames <- comps |>
  filter(!is.na(property_name))

comp_names <- comps_w_propnames |> 
  pull(property_name)

prop_names <- props |>
  filter(!is.na(property_name)) |>
  select(property_name) |>
  unlist()


comp_dists <- lapply(
  comp_names,function(x){
    
#    all_dists <- stringdist::stringsim(x,prop_names,
    # method="lcs") # another potential match algorithm
    all_dists <- stringdist::stringsim(x,prop_names,method="qgram",q=3)    
    closest_ind <- which(all_dists==max(all_dists))[1]    
    data.frame(
      closest=prop_names[closest_ind],
      closest_score=max(all_dists)
  )
  }
)
comp_dists <- do.call("bind_rows",comp_dists)

```

```{r}
comps_w_propnames <- comps_w_propnames |>
  ungroup() |>
  mutate(
    closest=pull(comp_dists,closest),
    closest_score=pull(comp_dists,closest_score),
    .after=property_name
  )
#View(comps_w_propnames |> arrange(desc(closest_score)))
```

Number of likely matches
```{r}
thr <- .6 # based on a high level review. this seems to be the point below which there are more nonmatches than matches
sum(comps_w_propnames$closest_score>thr)

```

