---
title: "connect-properties-complaints-location"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Complaints to Properties by Location

Match complaints to properties based on property name and address. This considers the 
complaints that were not previously matched by ID.
Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212241865138345?focus=true)

Inputs:
1. Complaints
  A. Unmatched by ID
  B. Duplicated by ID
  C. Missing ID
2. Cleaned Property file

Result of this script should be:

1. Complaints matched to properties
2. Unmatched Complaints

## INIT
```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
library(sf)
source("fns_dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260428" # date of file exports

```

## Read in data

### Complaints
```{r}
comps_unmatched_orig <- readRDS(paste(
  data_dir,"comps_unmatched_by_id.RDS",sep="/")) |>
  mutate(from="UNMATCHED",source="COMPS")

comps_miss_id_orig <- readRDS(paste(
  data_dir,"complaints_missing_id_dk.RDS",sep="/")) |>
  mutate(from="MISSING",source="COMPS")


comps_dupe_id_orig <- read_excel(paste(
  data_dir,"duplicate_complaints_review.xlsx",sep="/")) |>
  mutate(from="DUPE",source="COMPS")

# combine
comps <- bind_common_cols(
  comps_dupe_id_orig,comps_miss_id_orig,comps_unmatched_orig
)


```

### Properties
```{r}
props_id_orig <- readRDS(paste(data_dir,"properties_parcel_match_dk.RDS",sep="/")) |>
  mutate(source="PROPS")
props_inactive_orig <- readRDS(paste(data_dir,"properties_inactive_dontmatch_dk.RDS",sep="/")) |>
  mutate(source="PROPS")
props_unmatch_orig <- readRDS(paste(data_dir,"properties_address_match_dk.RDS",sep="/")) |>
  mutate(source="PROPS")

# combine
props <- bind_common_cols(
  props_id_orig,props_inactive_orig,props_unmatch_orig
)


```


## String Cleaning 
```{r}
comps <- comps |> clean_prop_names()
props <- props |> clean_prop_names()
  
```


## Initial Match Exploration
### Anti-join on property names
```{r}
comps_wname <- comps |> 
  filter(!is.na(property_name))
props_wname <- props |> 
  filter(!is.na(property_name))

name_disjoint <- comps_wname |> 
  anti_join(props_wname,by="property_name") |>
  select(property_name,source) |>
  count(property_name,source) |>
  arrange(property_name,source)

has_name_match <- comps_wname |> 
  semi_join(props_wname,by="property_name")


print(paste("Number of matching properties by prop_name:",nrow(has_name_match)))
```


### fuzzy match property names
```{r}
library(reclin2)
# TO DO

```

```{r}
library(wordcloud)
wordcloud(unique(comps$property_name),max.words = 75,random.order = FALSE)
```

### address merge

#### Uniqueness in properties
```{r}
dupe_props <- props |>
  count_dat(property_address) |>
  filter(n!=1)

props |>
  semi_join(dupe_props, by="property_address") 

```


```{r}
has_addr_match <- comps |>
  anti_join(has_name_match,by="property_name") |>
  semi_join(props, by="property_address") 
```

```{r}
missing_addr <- comps |>
  anti_join(has_name_match,by="property_name") |>
  anti_join(props, by="property_address") 
```


```{r}
nrow(missing_addr)
nrow(has_addr_match)
nrow(has_name_match)
nrow(missing_addr)+nrow(has_addr_match)+nrow(has_name_match)
nrow(comps)
```



## Secondary Matches
These are based on less precision than the previous stages and may need manual
confirmation

### Nearby addresses
Convert to sf
```{r}
comps_sf <- missing_addr |> 
  filter(!is.na(arcgis_location.x) & !is.na(property_address)) |>
  st_as_sf(coords=c("arcgis_location.x","arcgis_location.y")) |>
  select(complaint_id,property_name,property_address)

st_crs(comps_sf) <- 4326

props_sf <- props |>
  filter(!is.na(arcgis_location.x) & !is.na(property_address)) |>
  st_as_sf(coords=c("arcgis_location.x","arcgis_location.y")) |>
  select(property_id,property_name,property_address) |>
  rename_with(~ paste0(.x,"_prop"))
st_crs(props_sf) <- 4326

```

Spatial join on distance
```{r}
comps_near_sf <- st_join(comps_sf,props_sf,
                         join=st_is_within_distance,dist=50,
                         left=FALSE) |> 
  select(contains("_id"),contains("address"),contains("name"))
  
```

#### Review Potential Matches
Unique matches
```{r}
dist_match_unique <- comps_near_sf |>
  mutate(n_matches=n(),.by=complaint_id) |>
  filter(n_matches==1) 

dist_match_unique |>
  nrow()

dist_match_unique |>
  View()
```

Non-unique matches
```{r}
mult <- comps_near_sf |>
  mutate(n_matches=n(),.by=complaint_id) |>
  filter(n_matches>1) |>
  arrange(complaint_id, property_address, property_address_prop) 
  
mult |> nrow()  
mult |>  View()
```

Non-matches plus non-unique matches
```{r}
inconc_add <- missing_addr |>
  filter(!complaint_id %in% dist_match_unique$complaint_id) |>
#  filter(!complaint_id %in% comps_near_sf$complaint_id | 
#           complaint_id %in% mult$complaint_id) |>
  select(original_id,date_complaint_received,inspection_date,hud_complaint,
         complaint_outcome_comments,complainant_name,property_name,
         property_address,geo_address,property_address_orig,property_zip_code,
         arcgis_location.x,arcgis_location.y)  

nrow(inconc_add)
```

### Fuzzy Matching on property name
Just getting a rough count for now
```{r}
inconc_w_propnames <- inconc_add |>
  filter(!is.na(property_name))

comp_names <- inconc_w_propnames |>
  select(property_name) |>
  unlist()

prop_names <- props |>
  filter(!is.na(property_name)) |>
  select(property_name) |>
  unlist()


comp_dists <- lapply(
  comp_names,function(x){
    
#    all_dists <- stringdist::stringsim(x,prop_names,method="lcs")
    all_dists <- stringdist::stringsim(x,prop_names,method="qgram",q=3)    
    closest_ind <- which(all_dists==max(all_dists))[1]    
    data.frame(
      closest=prop_names[closest_ind],
      closest_score=max(all_dists)
  )
  }
)
comp_dists <- do.call("bind_rows",comp_dists)

```

```{r}
inconc_w_propnames <- inconc_w_propnames |>
  mutate(
    closest=comp_dists$closest,
    closest_score=comp_dists$closest_score,
    .after=property_name
  )
View(inconc_w_propnames |> arrange(desc(closest_score)))
```



Other things to check
Address ranges
fuzzy matching for property names
management office is the property address (probably for comps)
  needs hhrips review
  do they have a list of mgmt companies and addresses?
more?


Reviewing nonmatches
```{r}
missing_addr |>
  select(original_id,date_complaint_received,inspection_date,hud_complaint,
         complaint_outcome_comments,complainant_name,property_name,
         property_address,geo_address,property_address_orig,property_zip_code,
         arcgis_location.x,arcgis_location.y) |> 
  filter(date_complaint_received>=as.Date("2025-06-01")) |>
  View()
```


```{r}
props |>
  filter(str_detect(property_address,"DENVER")) |> arrange(property_address)
#  filter(str_detect(property_name,"FREN"))


#12509 E 53RD TER, KANSAS CITY, MISSOURI, 64133 | exists as 12509 E 53rd St
#7351 NW OLD TIFFANY SPRINGS RD, BUILDING 3, KANSAS CITY, MISSOURI, 64153 | nearby (7441 and others) but not exact match
# 4840 PASEO BLVD, KANSAS CITY, MISSOURI, 64110 | looks like part of UNIVERSITY MEADOWS
# 3200 DENVER AVE, KANSAS CITY, MISSOURI, 64128 | looks like part of RIDGEVIEW HEIGHTS

```


Nonmatches over time
```{r}
missing_addr |>
  ggplot(mapping=aes(x=rollback(date_complaint_received))) +
  geom_bar()



```


