---
title: "deduplicate-complaints-dk"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Deduplicate Complaints

Deduplicating the HHRIP complaint file. Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212019335801108?focus=true)

Result of this script should be:

1. A file with complaints that have distinct IDs
2. A file with complaints that are missing IDs
3. A file with complaints that have duplicate IDs

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(readxl)
library(here)

data_dir <- here("data")
date_str <- "20251208" # date of file exports

```

### Load Data
```{r}
comps_orig <- read_excel(paste0(data_dir,"/complaints_",date_str,".xlsx"))
```


## Review/Tidy Data Data
This will not be a full review and cleaning, but checking some high level aspects of the dataset

```{r}
comps <- comps_orig |> 
  janitor::clean_names()  
```


### Inspection 
15,339 rows and 68 columns
```{r}
str(comps)
```

### Column-specific summaries
#### Booleans
Distribution
```{r}
comps |>
  summarise(across(is.logical,mean)) |>
  t() 
```


#### Dates
Date missing, before program start (2018-08-01) or after today. 

Outputs counts of each, out of 15,339 rows. Based on this check it seems like a 
pretty small number of dates are outside of the range (largest are related to when
the complaint was received, which could potentially be before the actual start date).

There is quite a lot of missingness, some of which feel expected (events that
didn't occur), while others do not (e.g responded date).

```{r}
comps |>
  summarise(across(
    is.POSIXct, 
    list(
      miss = ~sum(is.na(.x)),
      before_start = ~sum(.x < as.Date("2018-08-01"),na.rm = TRUE),
      after_now = ~sum(.x > Sys.Date(), na.rm = TRUE)
    )
  )
) |>
  t()
```

#### Text
Set to all caps and trimmed
```{r}
comps <- comps |> 
  mutate(
    across(is.character, ~str_squish(str_to_upper(.x)))
  )
```


Missingness/duplicate values of text variables. 

Relevant for later on: there are 229 missing IDs (1.5%) and 534 duplicates (3.5%). 
Address is only missing in 63 records. There are quite a few duplicates in location info 
(properties, addresses, etc.). This could be expected since properties can have multiple complaints, 
and large properties can have very many complaints.

```{r}
comps |>
  summarise(across(
    is.character, 
    list(
      miss = ~sum(is.na(.x)),
      duplicates = ~sum(duplicated(.x,incomparables=NA))
    )
  )
) |>
  t()
```

### Variation among identifiers
Number of distinct property names within addresses
```{r}
# all addressXname combinations
addr_names <- comps |> 
  mutate(property_name=str_to_upper(property_name),
         property_address=str_to_upper(property_address)) |>
  count(property_address,property_name)  

# number of names for each address 
addr_names |>
  count(property_address,name="n_propnames") |>
  arrange(desc(n_propnames))

```

Number of distinct property names within addresses
```{r}
# number of addresses for each name
addr_names |>
  count(property_name,name="n_addresses") |>
  arrange(desc(n_addresses))
```

There's a huge amount of address variation for the same property name, and a decent amount of property name 
variation for the same address. The former is at least partially expected since large properties will have 
lots of addresses (and also there's tons of ways to write addresses). The latter is probably more due to data
entry issues.

Diferent 311 ID for the same property and date (note that this is a very rough check)

```{r}
comps <- comps |>
  mutate(loc_date=ifelse(
    !is.na(property_name), 
    paste(str_to_upper(property_name),as.character(date_complaint_received),sep="-"),
    paste(str_to_upper(property_address),as.character(date_complaint_received),sep="-")    
  ))

comps |>
  count(loc_date,x311_complaint_id) |>
  count(loc_date,name="n_compids") |>
  arrange(desc(n_compids))

```

```{r, include=FALSE}
# space for spot checks
comps |>
  filter(loc_date=="HUDSON POINTE APARTMENTS-2024-03-21")
```

Based on spot checks, it seems like some of these are from duplicate complaints, while others are from different
issues reported on the same property at the same day. We'll need to work with HHRIP to figure out how to handle these.

### 311 ID Data Quality 

```{r}
comps |>
  filter(str_detect(x311_complaint_id,"^[0-9]*$",negate=TRUE)) |>
  count(x311_complaint_id)
```

Looking at the ".1" (and .2,.3, etc) IDs

```{r}
comps |>
  filter(str_detect(x311_complaint_id,"\\.|-")) |>
  arrange(x311_complaint_id)

```


It looks like the nonnumeric characters in 311 IDs are from either:

1. A "suffix" on a duplicated record (987 vs 987.1)
2. The same ID was entered multiple times in the field, or text was accidentally entered after the ID
3. Very rarely, something that doesn't indicate the ID. Sometimes this is a date

To address these, going to set dates to NA and then update the ID to stop once a non-number is encountered
```{r}
comps <- comps |>
  mutate(
    original_id=x311_complaint_id,
    x311_complaint_id=ifelse(
      str_detect(original_id,"[0-9]\\/2"),
      NA,
      str_extract(original_id,"^[0-9]+")
  )
  )
```


```{r}
comps |>
  filter(original_id!=x311_complaint_id |
           (is.na(x311_complaint_id)&!is.na(original_id))) |>
  select(x311_complaint_id,original_id)
```



### Other Potential Checks
* Date logic (e.g. close before open)
* Valid addresses/ZIP Codes 

## Separate into necessary datasets
### Duplicate ID Exploration

Looking at potential indicators in addition to 311 ID

#### Comments indicating duplicates
This is a rough check based only on the exact word "duplicate"
```{r}
comps <- comps |>
  group_by(x311_complaint_id) |>
  mutate(n_ids = n()) |>
  ungroup() 

comps |>
  filter(n_ids > 1 & 
           (str_detect(str_to_upper(complaint_detail),"DUPLICATE") 
            | str_detect(str_to_upper(complaint_description2),"DUPLICATE")
            | str_detect(str_to_upper(original_id),"DUPLICATE"))
         ) |>
  arrange(desc(n_ids),x311_complaint_id)

comps |>
  filter(n_ids == 1 & 
           (str_detect(str_to_upper(complaint_detail),"DUPLICATE") 
            | str_detect(str_to_upper(complaint_description2),"DUPLICATE")
            | str_detect(str_to_upper(original_id),"DUPLICATE"))
         ) |>  
  arrange(desc(n_ids),x311_complaint_id)

```

There are some cases where the text indicates it is a duplicate, but there are no 
other records with that ID. This makes up <1% of the total so probably isn't a huge issue.
For now treating those as unique (especially since they could potentially tie to a parcel number/311 case).

#### Date, location, complainant name
```{r}
comps |>
  count(loc_date,complainant_name,x311_complaint_id,name="n_comps") |>
  arrange(desc(n_comps)) |>
  filter(n_comps>1)

comps |>
  count(loc_date,complainant_name,x311_complaint_id) |>
  count(loc_date,complainant_name,name="n_compids") |>  
  arrange(desc(n_compids)) |>
  filter(n_compids>1)


```

There are a small number of instances (though they represent a decently large number of complaints)
where the same person put in a complaint on the same day. In many but not all of these, they have 
the same complaint ID. For those with the same ID, they will be identified with the ID dedupe.
For the others, they may be uncaught duplicates. Small enough # that I'm going to leave them as is for now.

```{r, include=FALSE}
# space for spot checks
comps |>
  filter(loc_date=="11320 SYCAMORE TERR-2021-10-07" & 
           complainant_name=="MARIAH GORDON") |>
  glimpse()

```


### Separate Missing IDs
```{r}
comps_miss <- comps |>
  filter(is.na(x311_complaint_id))  

# remove from comps so that they don't get called duplicates
comps <- comps |>
  filter(!is.na(x311_complaint_id))    

```


### Separate Duplicate IDs from unique IDs
```{r}
comps_dupe <- comps |> filter(n_ids > 1)

comps_unique <- comps |> filter(n_ids == 1)

```


## Export

Row Counts

```{r}
nrow(comps_miss)
nrow(comps_dupe)
nrow(comps_unique)
```

Add exports
```{r}
saveRDS(comps_miss,paste(data_dir,"complaints_missing_id_dk.RDS",sep="/"))
saveRDS(comps_dupe,paste(data_dir,"complaints_duplicate_id_dk.RDS",sep="/"))
saveRDS(comps_unique,paste(data_dir,"complaints_unique_id_dk.RDS",sep="/"))
```

