---
title: "deduplicate-complaints-dk"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Deduplicate Complaints

Deduplicating the HHRIP complaint file. Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212019335801108?focus=true)

Result of this script should be:

1. A file with complaints that have distinct IDs
2. A file with complaints that are missing IDs
3. A file with complaints that have duplicate IDs

NOTE: Original pull/analysis was done on data from 12/08/2025 and some statistics
may be outdated

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
source("fns_dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260213" # date of file exports

```

### Load Data
```{r}
comps_orig <- read_excel(paste0(data_dir,"/complaints_",date_str,".xlsx"))
```


## Review/Tidy Data
This will not be a full review and cleaning, but checking some high level aspects of the dataset

```{r}
# renaming x311_complaint_id to just complaint_id for simplicity and alignment with inspections
comps <- comps_orig |> 
  janitor::clean_names() |>
  rename(complaint_id=x311_complaint_id)  
```



### Inspection 
15,339 rows and 68 columns
```{r}
str(comps)
```

### Column-specific summaries
#### Booleans
Distribution
```{r}
comps |>
  summarise(across(is.logical,mean)) |>
  t() 
```


#### Dates

HHRIP has noted that items with a complaint date in 2000 are known duplicates that
do not need to be migrated. Removing those, as well as other very early dates (prior
to program initiation)
```{r}
comps <- comps |> 
  filter(date_complaint_received>=as.Date("2018-01-01"))
```

Date missing, before program start (2018-08-01) or after today. 

Outputs counts of each, out of 15,339 rows. Based on this check it seems like a 
pretty small number of dates are outside of the range

There is quite a lot of missingness, some of which feel expected (events that
didn't occur), while others do not (e.g responded date).

```{r}
comps |>
  summarise(across(
    is.POSIXct, 
    list(
      miss = ~sum(is.na(.x)),
      before_start = ~sum(.x < as.Date("2018-08-01"),na.rm = TRUE),
      after_now = ~sum(.x > Sys.Date(), na.rm = TRUE)
    )
  )
) |>
  t()
```


#### Text
Set to all caps and trimmed
```{r}
comps <- comps |> clean_strs()
```


Missingness/duplicate values of text variables. 

Relevant for later on: there are 229 missing IDs (1.5%) and 534 records that are duplicated (3.5%. This does not count the first instance of the duplicated ID.). 
Address is only missing in 63 records. There are quite a few duplicates in location info 
(properties, addresses, etc.). This could be expected since properties can have multiple complaints, and large properties can have very many complaints.

```{r}
comps |>
  summarise(across(
    is.character, 
    list(
      miss = ~sum(is.na(.x)),
      duplicates = ~sum(duplicated(.x,incomparables=NA))
    )
  )
) |>
  t()
```


## 311 ID Data Quality 

### Non-numeric characters
```{r}
comps |>
  filter(str_detect(complaint_id,"^[0-9]*$",negate=TRUE)) |>
  count(complaint_id)
```

Looking at the ".1" (and .2,.3, etc) IDs

```{r}
comps |>
  filter(str_detect(complaint_id,"\\.|-")) |>
  arrange(complaint_id)

```


It looks like the nonnumeric characters in 311 IDs are from either:

1. A "suffix" on a duplicated record (987 vs 987.1)
2. The same ID was entered multiple times in the field, or text was accidentally entered after the ID
3. Very rarely, something that doesn't indicate the ID. Sometimes this is a date

~To address these, going to set dates to NA and then update the ID to stop once a non-number is encountered~

Per HHRIP, suffixes indicate an additional issue that was found after the inspector 
started working on a case so they should be kept. Going to create a "root ID" 
column without these for address matching but keep them for migration.
```{r}
comps <- comps |>
  mutate(
    original_id=complaint_id,
    complaint_id=ifelse(
      str_detect(original_id,"[0-9]\\/2"),
      NA,
      original_id
    ),
    root_id = str_extract(complaint_id,"^[0-9]+")
  ) |>
  relocate(original_id,root_id)
```


```{r}
comps |>
  filter(original_id!=complaint_id |
           (is.na(complaint_id)&!is.na(original_id))) |>
  select(complaint_id,original_id)
```

### Length

Check for non-standard ID lengths. Using Root ID, as that will be the primary matching ID.

```{r}
comps |> 
  mutate(id_length = nchar(root_id)) |>
  count(id_length)

# check varying lengths of those with small counts
comps |>
  mutate(id_length = nchar(root_id)) |>  
  filter(id_length<6) |>  
  #  filter(id_length==8|id_length==9) |>
  #filter(id_length>10) |>  
  select(contains("id"))
```

Small number of non-standard lengths. It looks like these are typos/other errors. 
Going to remove them for matching (though they could be useful for review later on).
```{r}
comps <- comps |>
  mutate(id_length = nchar(root_id)) |>  
  mutate(
    complaint_id=ifelse(!id_length %in% c(6,7,10), NA, complaint_id),
    root_id=ifelse(!id_length %in% c(6,7,10), NA, root_id)
  )

```


## Variation among identifiers

### Number of distinct property names within addresses
```{r}
# all addressXname combinations
addr_names <- comps |> 
  mutate(property_name=str_to_upper(property_name),
         property_address=str_to_upper(property_address)) |>
  count(property_address,property_name)  

# number of names for each address 
addr_names |> count_dat(property_address)

```

### Number of distinct addresses within property names
```{r}
# number of addresses for each name
addr_names |> 
  count_dat(property_name)
```

There's a huge amount of address variation for the same property name, and a decent amount of property name 
variation for the same address. The former is at least partially expected since large properties will have 
lots of addresses (and also there's tons of ways to write addresses). The latter is probably more due to data
entry issues.

### Diferent 311 ID for the same property name and date 

This could underestimate the issue as there is minimal address cleaning performed.

```{r}
comps <- comps |>
  mutate(loc_date=ifelse(
    !is.na(property_name), 
    paste(str_to_upper(property_name),as.character(date_complaint_received),sep="-"),
    paste(str_to_upper(property_address),as.character(date_complaint_received),sep="-")    
  ))

multCompid_sameDateloc <- comps |>
  filter(!str_detect(loc_date,"N\\/A") & !is.na(root_id)) |>
  count(loc_date,root_id) |>
  count(loc_date,name="n_compids") |> 
  filter(n_compids > 1) |>
  arrange(desc(n_compids))
multCompid_sameDateloc

```


Based on spot checks, it seems like some of these are from duplicate complaints, while others are from different
issues reported on the same property at the same day. We'll need to work with HHRIP to figure out how to handle these.

For example, it looks like complaint may have been received by MyCivic as well by HHRIP directly
```{r}
comps |>
  filter(loc_date=="103 W 51ST ST-2024-08-14") |>  
  select(complaint_id,complaint_recived_by,complaint_assigned_to,inspector_assigned_to,
         entered_in311on,date_complaint_received,
         complainant_name,complainant_phone,complaint_completed_date,no_violations_found,
         re_inspection_issued,complaint_detail)
```

Per HHRIP, if one of the complaint_received_by values is 311 and the other is a person,
those can likely be de-duplicated. We should keep them both for now for matching purposes,
but they don't need follow-ups.

```{r}
# flag records with one 311 complaint and one non-311 complaint, using 
# entries that contain "civic" or the number 3 as the indicator
comps <- comps |> 
  mutate(from_311=str_detect(complaint_recived_by,"3|CIVIC")) |>
  group_by(loc_date) |>
  mutate(diff_source=max(from_311)!=min(from_311))


comps |>
  inner_join(multCompid_sameDateloc) |>   
  filter(n_compids==2 & diff_source) |>
  select(loc_date,complaint_id,complaint_recived_by,from_311,n_compids) |> 
  arrange(desc(n_compids),loc_date)
```



Exporting for review
```{r}
lookup <- c(`Location/Date`="loc_date",`Number of Duplicate Locations/Dates`="n_locdate", 
            `Number of Duplicate IDs` = "n_compids",
            `At Least 1 Comp from 311 and 1 Not`="diff_source")
comps |>
  inner_join(multCompid_sameDateloc) |> 
  add_count(loc_date,name="n_locdate") |>
  filter(!(diff_source & n_compids==2)) |>
  arrange(loc_date,complaint_id) |>  
  select(loc_date,complaint_id,complaint_recived_by,diff_source,n_locdate,n_compids,
         complaint_assigned_to,inspector_assigned_to,
         entered_in311on,date_complaint_received,
         complainant_name,complainant_phone,complaint_completed_date,no_violations_found,
         re_inspection_issued,complaint_detail) |>
  rename(all_of(lookup)) |>
  rename_with(\(x) str_to_title(str_replace_all(x,"_"," "))) |>
  make_logicals_easier() |>
  write_xlsx(paste(outputs_dir,
                   "Complaints with Multiple IDs and the Same Complaint Info.xlsx",sep="/"))
```



## Other Potential Checks
* Date logic (e.g. close before open)
* Valid addresses/ZIP Codes 

## Separate into necessary datasets
### Duplicate ID Exploration

Looking at potential indicators in addition to 311 ID

#### Comments indicating duplicates
This is a rough check based only on the exact word "duplicate"
```{r}
comps <- comps |>
  group_by(complaint_id) |>
  mutate(n_ids = n()) |>
  ungroup() 

comps |>
  filter(n_ids > 1 & 
           (str_detect(str_to_upper(complaint_detail),"DUPLICATE") 
            | str_detect(str_to_upper(complaint_description2),"DUPLICATE")
            | str_detect(str_to_upper(original_id),"DUPLICATE"))
         ) |>
  arrange(desc(n_ids),complaint_id)

comps |>
  filter(n_ids == 1 & 
           (str_detect(str_to_upper(complaint_detail),"DUPLICATE") 
            | str_detect(str_to_upper(complaint_description2),"DUPLICATE")
            | str_detect(str_to_upper(original_id),"DUPLICATE"))
         ) |>  
  arrange(desc(n_ids),complaint_id)

```

There are some cases where the text indicates it is a duplicate, but there are no 
other records with that ID. This makes up <1% of the total so probably isn't a huge issue.
For now treating those as unique (especially since they could potentially tie to a parcel number/311 case).

#### Date, location, complainant name
```{r}
comps |>
  count(loc_date,complainant_name,root_id,name="n_comps") |>
  arrange(desc(n_comps)) |>
  filter(n_comps>1)

comps |>
  count(loc_date,complainant_name,root_id) |>
  count(loc_date,complainant_name,name="n_compids") |>  
  arrange(desc(n_compids)) |>
  filter(n_compids>1)


```


There are a small number of instances (though they represent a decently large number of complaints)
where the same person put in a complaint on the same day. In many but not all of these, they have 
the same complaint ID. For those with the same ID, they will be identified with the ID dedupe.
For the others, they may be uncaught duplicates. Small enough # that I'm going to leave them as is for now.

```{r, include=FALSE}
# space for spot checks
comps |>
  filter(loc_date=="11320 SYCAMORE TERR-2021-10-07" & 
           complainant_name=="MARIAH GORDON") |>
  glimpse()

```


### Separate Missing IDs
```{r}
comps_miss <- comps |>
  filter(is.na(complaint_id))  

# remove from comps so that they don't get called duplicates
comps <- comps |>
  filter(!is.na(complaint_id))    

```


### Separate Duplicate IDs from unique IDs

Note that there could still be duplicates here, because the same property can have a complaint on
the same day, but be assigned different IDs. Will keep those all for matching purposes.
```{r}
comps_dupe <- comps |> filter(n_ids > 1)

comps_unique <- comps |> filter(n_ids == 1)

```


## Export

Row Counts

```{r}
nrow(comps_miss)
nrow(comps_dupe)
nrow(comps_unique)
```

Add exports
```{r}
saveRDS(comps_miss,paste(data_dir,"complaints_missing_id_dk.RDS",sep="/"))
saveRDS(comps_dupe,paste(data_dir,"complaints_duplicate_id_dk.RDS",sep="/"))
saveRDS(comps_unique,paste(data_dir,"complaints_unique_id_dk.RDS",sep="/"))
```

