---
title: "deduplicate-complaints-dk"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review Duplicate Complaints

For complaints with duplicate IDs, matching . Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212241865138317?focus=true)

Result of this script should be:

1. A file with de-duplicated complaints when their IDs refer to the same complaint
2. A file with remaining duplicate IDs that seem to refer to different complaints,
which may need review


TO DO:
* Add more duplicate checks (e.g. could use geocoded address)
* Identify which complaint to keep when there are true duplicates

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
source("fns_dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260213" # date of file exports

set.seed(111)
```

### Load Data
```{r}
comps <- readRDS(paste(data_dir,"complaints_duplicate_id_dk.RDS",sep="/"))

comps <- comps |>
  group_by(complaint_id) |>
  arrange(complaint_id) |>
  mutate(comp_num=row_number())
```


## Flag true duplicates

### Date of complaint
```{r}
day_mark <- 2
comps <- comps |>
  mutate(
    same_date = date_complaint_received ==
      lag(date_complaint_received, 1, order_by = comp_num),
    close_date = as.numeric(
      difftime(date_complaint_received,
        lag(date_complaint_received, 1, order_by = comp_num),
        units = "days"
      )
    ) <= day_mark
  )

```

### Location of complaint
```{r}
comps <- comps |> 
  mutate(
    last_propname=lag(property_name, 1, order_by = comp_num),
    last_addr=lag(property_address, 1, order_by = comp_num),    
  ) |>
  mutate(
    same_prop=property_name==last_propname,
    same_addr=property_address==last_addr,
    same_loc=case_when(
      comp_num==1 ~ NA,
      is.na(same_prop) & is.na(same_addr) ~ FALSE,
      same_prop | same_addr ~ TRUE,
      TRUE ~ FALSE
    )
  )
    
```

Space for ad hoc reviews
```{r}
#comps |>
#  filter(complaint_id=="1010640") |>  
  #filter(complaint_id=="1747318") |>
  #filter(complaint_id=="1063473") |>
#  select(property_name,property_address,last_propname,last_addr,same_prop,same_addr,same_loc)
```


### Complainant name

### Similarity in notes



### Flag duplicates
NOTE: Commented items are for exact date matches
```{r}
comps <- comps |>
  mutate(same_as_above=close_date & same_loc)
#  mutate(same_as_above=same_date & same_loc)
```



## De-duplicate
Several potential strategies:
1. Status (opt for closed)
2. Recency (opt for most recent date...though not sure which)
3. More populated fields?

Separate problematic duplicates (e.g. those that seem to refer to different properties)
from those that can be safely de-duplicated
```{r}
comps_with_dupes <- comps |>
  mutate(has_problematic_dupes=!as.logical(min(same_as_above,na.rm = TRUE))) |>
  filter(has_problematic_dupes)

dupe_comp_ids <- comps_with_dupes$complaint_id
```


```{r}
comps_deduped <- comps |>
  filter(!complaint_id %in% dupe_comp_ids) |>
  distinct(complaint_id,
#           date_complaint_received,
           property_name,
           .keep_all = TRUE) |>
  distinct(complaint_id,
#           date_complaint_received,
           property_address,
           .keep_all = TRUE)
```

## Export
```{r}
write_xlsx(comps_with_dupes,paste(data_dir,"duplicate_complaints_review.xlsx",sep="/"))
saveRDS(comps_deduped,paste(data_dir,"complaints_deduplicated_dk.RDS",sep="/"))
```

