---
title: "connect-properties-complaints-location"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Complaints to Properties by Location

Match complaints to properties based on property name and address. This considers the 
complaints that were not previously matched by ID.
Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212241865138345?focus=true)

Inputs:
1. Complaints
  A. Unmatched by ID
  B. Duplicated by ID
  C. Missing ID
2. Cleaned Property file

Result of this script should be:

1. Complaints matched to properties
2. Unmatched Complaints

## INIT
```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(here)
source("fns_dk.R")

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260428" # date of file exports

```

## Read in data

### Complaints
```{r}
comps_unmatched_orig <- readRDS(paste(
  data_dir,"comps_unmatched_by_id.RDS",sep="/")) |>
  mutate(from="UNMATCHED",source="COMPS")

comps_miss_id_orig <- readRDS(paste(
  data_dir,"complaints_missing_id_dk.RDS",sep="/")) |>
  mutate(from="MISSING",source="COMPS")


comps_dupe_id_orig <- read_excel(paste(
  data_dir,"duplicate_complaints_review.xlsx",sep="/")) |>
  mutate(from="DUPE",source="COMPS")

# combine
comps <- bind_common_cols(
  comps_dupe_id_orig,comps_miss_id_orig,comps_unmatched_orig
)


```

### Properties
```{r}
props_id_orig <- readRDS(paste(data_dir,"properties_parcel_match_dk.RDS",sep="/")) |>
  mutate(source="PROPS")
props_inactive_orig <- readRDS(paste(data_dir,"properties_inactive_dontmatch_dk.RDS",sep="/")) |>
  mutate(source="PROPS")
props_unmatch_orig <- readRDS(paste(data_dir,"properties_address_match_dk.RDS",sep="/")) |>
  mutate(source="PROPS")

# combine
props <- bind_common_cols(
  props_id_orig,props_inactive_orig,props_unmatch_orig
)


```


## String Cleaning 
```{r}
comps <- comps |> clean_prop_names()
props <- props |> clean_prop_names()
  
```


## Initial Match Exploration
### Anti-join on property names
```{r}
comps_wname <- comps |> 
  filter(!is.na(property_name))
props_wname <- props |> 
  filter(!is.na(property_name))

name_disjoint <- comps_wname |> 
  anti_join(props_wname,by="property_name") |>
  select(property_name,source) |>
  count(property_name,source) |>
  arrange(property_name,source)

has_name_match <- comps_wname |> 
  semi_join(props_wname,by="property_name")


print(paste("Number of matching properties by prop_name:",nrow(has_name_match)))
```


### fuzzy match property names
```{r}
library(reclin2)
# TO DO

```

```{r}
library(wordcloud)
wordcloud(unique(comps$property_name),max.words = 75,random.order = FALSE)
```

### address merge

#### Uniqueness in properties
```{r}
dupe_props <- props |>
  count_dat(property_address) |>
  filter(n!=1)

props |>
  semi_join(dupe_props, by="property_address") 

```


```{r}
has_addr_match <- comps |>
  anti_join(has_name_match,by="property_name") |>
  semi_join(props, by="property_address") 
```

```{r}
missing_addr <- comps |>
  anti_join(has_name_match,by="property_name") |>
  anti_join(props, by="property_address") 
```


```{r}
nrow(missing_addr)
nrow(has_addr_match)
nrow(has_name_match)
nrow(missing_addr)+nrow(has_addr_match)+nrow(has_name_match)
nrow(comps)
```


Other things to check
Address ranges
fuzzy matching for complaints
management office is the property address (probably for comps)
  needs hhrips review
  do they have a list of mgmt companies and addresses?
more?


