---
title: "connect-properties-311"
author: "Dylan Knaggs"
date: "2025-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Match Properties with Complaints by ID

Combine Complaints with the properties dataset [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212050138239287?focus=true)


## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(tidygeocoder)
library(readxl)
library(janitor)
library(here)


data_dir <- here("data")
date_str <- "20251208" # date of file exports

source(paste(here("scripts"),"fns_dk.R",sep="//"))

```


## Load Data

### Properties
```{r}
props <- readRDS(paste(data_dir,"properties_parcel_match_dk.RDS",sep="/")) |>
  mutate(props=1)
```

### Parcels
```{r}
parcels <- read_csv(paste(data_dir,"parcels_20251210.csv",
                   sep="/")) |>
  clean_names() |>
  mutate(parcels=1)

```


### 311 Calls
```{r}
calls_new <- read_xlsx(paste(data_dir,"311-cases.xlsx",
#calls_new <- read_csv(paste(data_dir,"311_calls_post_mar21_20251210.csv", #original download
#calls_new <- read_csv(paste(data_dir,"hhrip-complaints.csv", # updated download by dept
                   sep="/")) |>
  clean_names() |>
  mutate(calls=1)

calls_old <- read_csv(paste(data_dir,"311_calls_pre_mar21_20251210.csv",
                   sep="/")) |>
  clean_names() |>
  filter(work_group=="Health-Healthy Homes-") |>
  mutate(calls=1)

```

### Access Complaints
```{r}
comps <- bind_rows(
  readRDS(paste(data_dir,"complaints_unique_id_dk.RDS",sep="/")),
  readRDS(paste(data_dir,"complaints_deduplicated_dk.RDS",sep="/"))
) |>
  mutate(comps=1)
```



## Align Columns

### 311 Calls

Column types
```{r}
calls_new <- calls_new |>
  mutate(
    id=as.character(id),
    pin=as.character(pin),
    open_date=as.Date(open_date,"%m/%d/%Y"),    
    resolved_date=as.Date(resolved_date,"%m/%d/%Y"),    
  )

calls_old <- calls_old |>
  mutate(
    case_id=as.character(case_id),
    parcel_id_no=as.character(parcel_id_no),
    creation_date=as.Date(creation_date,"%m/%d/%Y"),    
    closed_date=as.Date(closed_date,"%m/%d/%Y"),    
  )

```

Column names (and bind new+old calls)
```{r}
calls <- bind_rows(
  calls_new |> 
    select(id,pin,status,open_date,resolved_date,
           address,latitude,longitude,open_date,calls)|>
    rename(property_address=address,complaint_id=id),
  
  calls_old |> 
    select(case_id,parcel_id_no,status,creation_date,closed_date,
           street_address,latitude,longitude,calls) |>
    rename(complaint_id=case_id,pin=parcel_id_no,open_date=creation_date,
           resolved_date=closed_date,property_address=street_address)
  )
  
```


### parcels
```{r}
parcels <- parcels |>
  select(kivapin,apn,address,landusecode,parcels) |>
  rename(pin=kivapin,parcel_num=apn,property_address=address)

```

```{r}
parcels <- parcels |>
  mutate(pin=as.character(pin))

```


## Check Uniqueness 

### Properties
```{r}
props |>
  mutate(parcel_ct=n(),.by=parcel_num) |>
  filter(parcel_ct>1) |>
  arrange(parcel_num)
```

NOTE:
there are a small number of properties with duplicate parcel numbers. These
should be handled in deduplicate-properties.Rmd, but going to remove here for now,
keeping the most recent.

```{r}
props <- props |>
  arrange(desc(application_date)) |>
  distinct(parcel_num,.keep_all = TRUE)
```


### Parcels
```{r}
parcels |> 
  count_dat(pin) |>
  filter(n>1)

parcels |> 
  count_dat(parcel_num) |>
  filter(n>1)
```


There are also a small number of parcels with duplicate parcel numbers or 
pins. Deduplicating arbitrarily for now.

```{r}
parcels <- parcels |> 
  distinct(parcel_num,.keep_all = TRUE) |>
  distinct(pin,.keep_all = TRUE)  
```


### 311 Calls
```{r}
calls |> 
  count_dat(complaint_id) |>
  filter(n>1)
```


### Access Complaints

```{r}
comps |> 
  count_dat(complaint_id) |>
  filter(n>1)
```


## Join datasets

### Access complaints to calls
```{r}
comps_311 <- comps |>
  left_join(calls, by="complaint_id",suffix=c("","_311"),
            relationship="one-to-one") 

unmatched_to_311 <- comps_311 |>
  filter(is.na(calls))

comps_311 <- comps_311 |>
  filter(!is.na(calls))

```

### Include parcels
```{r}
comps_par <- comps_311 |>
  left_join(parcels,by="pin",suffix=c("","_par"),
            relationship = "many-to-one")

unmatched_to_par <- comps_par |>
  filter(is.na(parcels))

comps_par <- comps_par |>
  filter(!is.na(parcels))

```

### Include properties
```{r}
comps_prop <- comps_par |>
  left_join(props,by="parcel_num",suffix=c("","_prop"),
            relationship = "many-to-one")

unmatched_to_prop <- comps_prop |>
  filter(is.na(props))

comps_prop <- comps_prop |>
  filter(!is.na(props))


```


## Review
Confirm row counts and check number of matches by stage 
```{r}
all_unmatched <- bind_rows(
  unmatched_to_311,
  unmatched_to_par,
  unmatched_to_prop
)
# Unmatched at each stage
nrow(all_unmatched)
nrow(unmatched_to_311)
nrow(unmatched_to_par)

# confirm all properties are accounted for
nrow(all_unmatched)+nrow(comps_prop)
nrow(comps)
```

### Review nonmatches
Biggest issues occurs from parcel to property match. Seeing if there are trends in mismatches over time
```{r}
unmatched_to_prop |> 
  mutate(month=floor_date(date_complaint_received,unit="month")) |>
  count(month) |>
  print(max=500)
  #print(n=Inf)

```


Spot checks
```{r}
props_all <- read_csv(paste(data_dir,"active_props_20251208.csv",sep="/")) |>
  clean_strs() |>
  clean_names()


```
```{r}
all_unmatched |>
  filter(parcels==1) |>
  select(complaint_id,property_address,property_name,date_complaint_received,
         parcel_num,pin,comps,calls,parcels) |>
  arrange(desc(date_complaint_received))


```


```{r}
props_all |>
#  filter(str_detect(owner_name,"EPD")) |>
  filter(str_detect(property_address,"63RD")) |>
#  filter(str_detect(parcel_num,"JA28840021301000000")) |>
  arrange(property_address) |>
  select(property_address,property_name,owner_name,parcel_num)


```


Reasons why missingness could occur here:
1. duplicate parcel, though in many cases one of these is active and the other is inactive
  Could do an inequality join based on complaint/registration date (though not sure complaint date
  will always be after registration date)
2. Missing parcel numbers
3. Might not be registered? ( 608 E 43rd)
4. Different addresses/parcels (RIDGEVIEW HEIGHTS APARTMENTS)

Next Steps:
Refresh data sources
Export remaining unmatched (and unmatchable) complaints
Export cleaned parcels
Property name/address match
