---
title: "deduplicate-properties-dk"
author: "Dylan Knaggs"
date: "2025-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Deduplicate Properties

Deduplicating the HHRIP properties file. Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212050138239285?focus=true)

This is more of a high level quality check than a full data deduplication process, as those were cleaned prior to being uploaded to the MSSQL db. Assuming a reasonable degree of uniqueness in the properties data we will not change the data beyond adding a flag for potential duplicates

## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(tidygeocoder)
library(readxl)
library(here)


data_dir <- here("data")
date_str <- "20251208" # date of file exports

source(paste(here("scripts"),"fns_dk.R",sep="//"))

```


### Load Data
```{r}
active_orig <- read_csv(paste0(data_dir,"/active_props_",date_str,".csv"))
inactive_orig <- read_csv(paste0(data_dir,"/inactive_props_",date_str,".csv"))
```


## Review/Tidy Data Data
This will not be a full review and cleaning, but checking some high level aspects of the dataset

```{r}
active <- active_orig |> 
  janitor::clean_names()  
inactive <- inactive_orig |> 
  janitor::clean_names()

```


### Inspection 
18,338 rows and 17 columns
```{r}
str(active)
```
5,689 rows and 10 columns
```{r}
str(inactive)
```
### Keep key columns and bind active/inactive
```{r}
inactive <- inactive |> mutate(from="inactive")

props <- bind_rows(
  active |>
    mutate(from="active") |>
    select(all_of(
      c(names(inactive),"owner_id","owner_name")
    )), 
  inactive
)

if(exists("active")){
  rm(active,inactive)
}
```

### Capitalize text columns 
```{r}
props <- clean_strs(props)
```



### set ID/ZIP variables to character
```{r}
props <- props |> mutate(across(c(property_id,owner_id,property_zip), ~as.character(.x)))

```

### Identify invalid parcel IDs
Manual checks showed that some Parcel IDs weren't valid. These should be set to NA for address matching. 
They can be variable lengths, but based on distribution, think they should be at least 14 and lnot more than 24 characters
```{r}
props <- props |> mutate(parcel_len = nchar(parcel_num))
props |>
  count_dat(parcel_len) |>
  arrange(parcel_len)

```

```{r}
props <- props |>
  mutate(parcel_num_orig=parcel_num) |>
  mutate(parcel_num=ifelse(parcel_len < 14 | parcel_len >24 ,NA,parcel_num)) |>
  select(-parcel_len)
```



### Check missingness of variables
```{r}
col_missingness(props)
```

A couple of notes:
1. Owner information is missing from all inactive properties. This may or may not affect things down the line.
2. lat and long columns are present but largely missing
3. Parcel number is missing in 13% of cases. This translates to at least 3,100 properties that will need to be fuzzy matched 


### Remove redundant inactives
From later analysis, it was found that there are duplicates that arise because. Setting aside inactive properties that seem to be duplicates based on address. 

```{r}
dupe_addrs <- props |> 
  count_dat(property_address) |>
  filter(n>1)

inactive_dupes <- props |>
  filter(from=="INACTIVE" 
         & property_address %in% dupe_addrs$property_address
  )

props <- props |>
  filter(!(from=="INACTIVE" 
         & property_address %in% dupe_addrs$property_address)
  )

# should keep the same total number
print(nrow(inactive_dupes)+nrow(props))
print(nrow(active_orig)+nrow(inactive_orig))

```

## Flagging Duplicates
Using a couple of different strategies to check various results

### Same property ID
These _should_ be unique
```{r}
props |> count_dat(property_id)

```


### Same parcel ID
```{r}

props |> count_dat(parcel_num)


props <- props |>
  group_by(parcel_num) |>
  mutate(
    n_parcels = ifelse(is.na(parcel_num),NA,n()),
    dup_parcels = ifelse(n_parcels>1,TRUE,FALSE)
  ) |>
  ungroup()

```
About 4% (1,024) properties have the same parcel as one other. About 1% (213 have multiple duplicates).

```{r}
props |>
  filter(dup_parcels) |>
  arrange(parcel_num) |>
  select(property_id,parcel_num,property_address,property_name,
         owner_id,owner_name,application_date)
```
From manual inspections, there seems to be 3 sources of duplicates:
1. Genuine duplicates: these can be flagged/removed (maybe based on parcel,address,application date?)
2. Properties that were rentals, went inactive, then were applied for ownership again: probably best to have a separate flag for these (same parcel and address, different application date)
3. Multiple addresses on the same parcels: these probably need to be address-matched


### Same address

Eventually want to geocode addresses to ensure more consistency but this will take a very long time 
with single address geocoding. Eventually could use something like this
```{r,eval=FALSE}
#props_geo <- props |> 
#  mutate(addr_geocode=paste(property_address,"KANSAS CITY, MO",property_zip)) |>
#  select(property_id, addr_geocode) |>
#  tail(50) |>
#  geocode(addr_geocode,method="arcgis",full_results=TRUE,progress_bar=TRUE)

```

NOTE: This chunk has failed to run intermittently. I think it's related to the session rather than the code, but need to keep an eye out.
```{r}

props |> count_dat(property_address)


props <- props |>
  group_by(property_address) |>
  mutate(
    n_addrs = ifelse(is.na(property_address),NA,n()),
    dup_addrs = ifelse(n_addrs>1,TRUE,FALSE)
  ) |>
  ungroup()

```

### Property Names
```{r}
props |> 
  count_dat(property_name) |>
  arrange(property_name)

```
Observations: 
1. Many property names are the same as the address
2. There's a decent amount of spelling variation, so some sort of strdist may be necessary


## Tagging Duplicates
### Same parcel, address, date

These are likely to be the same property in the same timeframe

Eventually want to geocode addresses to ensure more consistency but this will take a very long time 
with single address geocoding. Could do batch geocoding, but need to do that outside R/learn how to do it within R. For now using exact matches (could also consider strdist)

```{r}
all_dupe <- props |> 
  count_dat_mult(c("parcel_num","property_address","application_date")) |>
  filter(n>1 & !is.na(parcel_num))
all_dupe
```
Only one record here.

### Same parcel, address, different date

These are likely to be the same property but with different periods of being rentals (or potentially a duplicate with different dates entered)
```{r}
addr_parcel_dupe <- props |>
  count_dat_mult(c("parcel_num","property_address")) |>
  filter(n>1 & !is.na(parcel_num)) |>
  anti_join(all_dupe,by=c("parcel_num","property_address"))
addr_parcel_dupe

props |> 
  inner_join(addr_parcel_dupe) |>
  select(parcel_num,property_address,application_date,owner_name) |>
  arrange(parcel_num,application_date)

```
Just based on the dates/owners, some of these could be duplicates while others could be change of ownerships. Given that there are only 50 here and 1 in the set with identical dates, I'm just going to tag everything with the same parcel/address.


### Same parcel, different address
These are likely to be multiple rental properties on a single parcel
```{r}
one_par_mult_adds <- props |>
  count(parcel_num,property_address) |>
  count(parcel_num,name="n_addrs_parcel") |>  
  arrange(desc(n_addrs_parcel)) |>
  filter(n_addrs_parcel>1)

props |> 
  inner_join(one_par_mult_adds) |>
  select(parcel_num,property_address,application_date,owner_name) |>
  arrange(parcel_num,application_date)

```
From manual review, that generally seems to be the case. There are some situations where it seems like a duplicate with the address spelled incorrectly. Additional address cleaning could potentially help distinguish these cases.


### One address, multiple parcels
Also checking the inverse: a single address with multiple parcel IDs

```{r}
one_addr_mult_pars <- props |>
  count(property_address,parcel_num) |>
  count(property_address,name="n_parcels_addr") |>  
  arrange(desc(n_parcels_addr)) |>
  filter(n_parcels_addr>1)
one_addr_mult_pars

props |> 
  inner_join(one_addr_mult_pars) |>
  select(property_address,parcel_num,application_date,owner_name) |>
  arrange(property_address,application_date)


```

These seem to be duplicates, and it seems to happen because a property has two parcel IDs. Need to understand why this happens, but for now will keep all these records.

## Final tagging and exports

### Export missing parcels
Separating out all records with missing parcel IDs or multiple addresses per parcel, since these will need to be address-matched regardless of the situation. Note that it is possible that a duplicate, though address duplicates seem quite rare.
```{r}
# could maybe do this via joins
incomps <- props |>
  filter(is.na(parcel_num) | parcel_num %in% one_par_mult_adds$parcel_num)


props <- props |>
  filter(!parcel_num %in% one_par_mult_adds$parcel_num)


```

Ensuring everything got kept
```{r}

print(nrow(incomps)+nrow(props)+nrow(inactive_dupes))
print(nrow(active_orig)+nrow(inactive_orig))

```

Adding final tags
NOTE: There may be others that need to be included.
```{r}
props <- props |>
  mutate(one_addr_mult_pars=ifelse(property_address %in% one_addr_mult_pars,
         TRUE,FALSE)
  )
```


### Exporting
```{r}
saveRDS(inactive_dupes,paste(data_dir,"properties_inactive_dontmatch_dk.RDS",sep="/"))
saveRDS(incomps,paste(data_dir,"properties_address_match_dk.RDS",sep="/"))
saveRDS(props,paste(data_dir,"properties_parcel_match_dk.RDS",sep="/"))
```
