---
title: "deduplicate-complaints-dk"
author: "Dylan Knaggs"
date: "2025-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Harmonize Inspections

ORIGINAL ASSIGNMENT TEXT:
Harmonizing 311 IDs within HHRIP complaints. Project task is available [here] (https://app.asana.com/1/39550735975676/project/1211786938939379/task/1212241844370491?focus=true)

Result of this script should be:
1. A file with series of inspections that refer to the same location and have the same ID
2. A file with inspections that have potential ID issues (e.g. same ID but different inspection location)

I've ended up not doing this yet. I'll let this serve as a basic cleaning/exploration
script but there's a lot of different directions we could take the actual harmonization. I think it would be helpful to see how the merge with complaints looks
before we make too many changes to inspections. If we just manage to clean complaint
addresses (since there's only one per complaint) and have sufficient confidence that the IDs line up across complaints and
inspections, we don't need to repeat that work.


## Initialize

### Packages and Constants 
```{r init}
library(tidyverse)
library(readxl)
library(writexl)
library(stringdist)
library(here)

data_dir <- here("data")
outputs_dir <- here("outputs")
date_str <- "20260213" # date of file exports

source("fns_dk.R")

```

NOTE: Original pull/analysis was done on data from 12/08/2025 and some statistics
may be outdated


### Load Data
```{r}
col_types <- c("text",rep("guess",9),"text",rep("guess",54),"text") 

insps_orig <- read_excel(
  paste0(data_dir,"/inspection_history_",date_str,".xlsx"),
  col_types = col_types)
```


## Review/Tidy Data
This will not be a full review and cleaning, but checking some high level aspects of the dataset

```{r}
insps <- insps_orig |> 
  janitor::clean_names()  
```


### Inspection 
19,186 rows and 66 columns
```{r}
str(insps)
```

### Column-specific summaries
#### Booleans
Distribution
```{r}
insps |>
  summarise(across(is.logical,mean)) |>
  t() 
```


#### Dates

Date Entered can't be read as a date, explicitly casting it to one.
```{r}
insps <- insps |>
  mutate(date_entered=as.POSIXct(as.Date(date_entered,"%m/%d/%Y")))

```


Date missing, before program start (2018-08-01) or after today. 

Very small numbers of missingness/out of range values

```{r}
insps |>
  summarise(across(
    is.POSIXct, 
    list(
      miss = ~sum(is.na(.x)),
      before_start = ~sum(.x < as.Date("2018-08-01"),na.rm = TRUE),
      after_now = ~sum(.x > Sys.Date(), na.rm = TRUE)
    )
  )
) |>
  t()
```

Per HHRIP, inspections from 2000 are duplicates. Removing those, as well as other very early dates (prior to program initiation)
```{r}
insps <- insps |> 
  filter(inspection_date>=as.Date("2018-01-01"))
```


#### Text
Set to all caps and trimmed
```{r}
insps <- insps |> clean_strs()
```


Missingness/duplicate values of text variables. 

Most key fields have very low missingness (17 for complaint ID, 
17 for address, 52 for ZIP Code). 

Maybe notable that permit is almost always missing, but that isn't needed
for the analysis.

```{r}
insps |>
  summarise(across(
    is.character, 
    list(
      miss = ~sum(is.na(.x)),
      duplicates = ~sum(duplicated(.x,incomparables=NA))
    )
  )
) |>
  t()
```

There are a lot of duplicate values in inspections. Checking those out
```{r}
insps |>
  count_dat(complaint_id)
```
Some records do have a lot of inspections, but it doesn't seem unreasonable over 
a lot of years, and none of the complaint IDs look fake ("none" or something like that).


#### ZIP Codes
Checking ZIP Code quality
```{r}
insps |>
  filter(!str_detect(property_zip_code,"^[0-9][0-9][0-9][0-9][0-9]$")) |>
  select(property_zip_code)
```


Generally very few issues. Removing final characters for others that work. 
Note: this regex probably doesn't work in all situations but looks to be 
what is needed here.
```{r}
insps <- insps |>
  mutate(property_zip_code=str_extract(property_zip_code,"^[0-9]*")) 
```


## Data cleaning
### Number of addresses within complaint ID
```{r}
# all complaint idXaddress combinations
id_addrs <- insps |> 
  count_dat_mult(c("complaint_id","property_address"))

id_addrs

# number of names for each address 
mult_addrs <- id_addrs |>
  count_dat(complaint_id) |>
  filter(n>1) |>
  rename(n_addrs=n) 
mult_addrs

```

### Second check with geocoded addresses

Note that this excludes addresses that couldnt be geocoded

Note that inspection_addrs_geocoded.csv was generated in ArcGIS Pro. 
```{r}
insps_geo <- read_csv(paste0(data_dir,"/inspection_addrs_geocoded.csv")) |>
  filter(!is.na(Match_addr)) |>
  mutate(USER_property_zip_code=as.character(USER_property_zip_code))

insps_check <- insps |>
  inner_join(insps_geo,by=c("property_address"="USER_property_address",
                            "property_zip_code"="USER_property_zip_code"))

```

##### Original addrs
```{r}
# all complaint idXaddress combinations
id_addrs1 <- insps_check |> 
  count_dat_mult(c("complaint_id","property_address"))

id_addrs1

# number of addresses within each complaint_id
id_addrs1 |>
  count_dat(complaint_id) |>
  filter(n>1)

```


#### Geocoded addrs
```{r}
# all complaint idXaddress combinations
id_addrs1 <- insps_check |> 
  count_dat_mult(c("complaint_id","Match_addr"))

id_addrs1

# number of addresses within each complaint_id
mult_addrs1 <- id_addrs1 |>
  count_dat(complaint_id) |>
  filter(n>1) |>
  rename(n_addrs_geo=n)

```



```{r}
# space for spot checks
insps_check |>
  filter(complaint_id=="2276911") |>
  select(complaint_id,property_name, property_address,Match_addr)
```
Still not perfect alignment but this improved things drastically. Looks like most of the differences are from large properties with multiple addresses, or more substantial typos.

### Number of property names within complaint ID
```{r}
# all complaint idXproperty name combinations
# interested in seeing with and without NAs 

id_names <- insps |> 
  mutate(property_name=ifelse(is.na(property_name),"NONE",property_name)) |>
  count_dat_mult(c("complaint_id","property_name"))

id_names |> arrange(property_name)
id_names |> arrange(property_name) |> filter(n>1)

# number of names for each address 
# including NAs as separate entries
mult_names <- id_names |>
  count_dat(complaint_id) |>
  arrange(complaint_id) |>
  filter(n>1) |>
  rename(n_names=n) 
mult_names
  

# excluding NAs
id_names |>
  filter(property_name!="NONE") |>
  arrange(complaint_id) |>  
  count_dat(complaint_id) |>
  filter(n>1)

```

About 1,000 complaints have inspections with more than 1 (non-NA) property name.
From spot checks these largely due to typos/format differences.


### Property name AND address within complaint ID
```{r}
mult_names_addrs <- mult_names |> 
  inner_join(mult_addrs) |>
  inner_join(mult_addrs1)
mult_names_addrs

```


### 311 ID Data Quality
```{r}
insps |>
  filter(str_detect(complaint_id,"^[0-9]*$",negate=TRUE)) |>
  count(complaint_id)
```

Similar issues with suffixes as in complaints
~To address these, going to set dates to NA and then update the ID to stop once a non-number is encountered~

Per HHRIP, suffixes indicate an additional issue that was found after the inspector 
started working on a case so they should be kept. Going to create a "root ID" 
column without these for address matching but keep them for migration.

```{r}
insps <- insps |>
  mutate(
    original_id=complaint_id,
    complaint_id=ifelse(
      str_detect(complaint_id,"[0-9]\\/2"),
      NA,
      complaint_id
    ),
    root_id = str_extract(complaint_id,"^[0-9]+")
  ) |>
  relocate(original_id,root_id,complaint_id)

```



### Duplicates
#### Words that indicate duplicates
```{r}
insps |>
  filter(str_detect(notes,"DUPLIC"))

```

Very few records, dropping.
```{r}
insps <- insps |> filter(!str_detect(notes,"DUPLIC") | is.na(notes))
```



#### Same complaint id and date
Also considering inspection type (those with the same type are more likely to be duplicates)

```{r}
insps <- insps |>
  add_count(complaint_id,inspection_date,name="n_ids") |>
  add_count(complaint_id,inspection_date,inspection_type,name="n_ids_inspTypes")  

# overview of duplicates
insps |>
  count(n_ids)

# duplicate insp but different type (likely process-based) 
insps |>
  filter(n_ids != n_ids_inspTypes)

# duplicate insp and same type (more likely duplicate)
insps |>
  filter(n_ids>1 & n_ids == n_ids_inspTypes)

```

There are a substantial number of records with the same ID on the same day. 
It seems like this could plausibly happen in a few cases (large properties, same-day reinspections, desk approvals, etc) but working with HHRIP to understand.

If all violation information and outputs are the same, just deduplicating,
then recalculating ID counts
```{r}
insps |>
  count()

insps |>
  distinct(complaint_id,inspection_date) |>
  count()


insps |>
  distinct(complaint_id,inspection_date,re_inspection_required,pick(starts_with("x")))

insps <- insps |>
  distinct(complaint_id,inspection_date,re_inspection_required,
           pick(starts_with("x")),
           .keep_all = TRUE)


```

```{r}
insps <- insps |>
  add_count(complaint_id,inspection_date,name="n_ids") |>
  add_count(complaint_id,inspection_date,inspection_type,name="n_ids_inspTypes")  
```


For the others, counting the number of violations in common
```{r}
insps <- insps |> 
group_by(complaint_id,inspection_date) |>
  mutate(
    n_common_viols=sum(
      across(starts_with("x"),~sum(min(.x)==max(.x)))
    ),
    same_reinspection_status = 
      max(re_inspection_required)==min(re_inspection_required)
  ) |>
  ungroup()
```

Almost all duplicates have the majority of violations in common and many 
have >40 out of 50 in common
```{r}
insps |>
  filter(n_ids>1) |>
  count(n_common_viols)

```


Exporting those duplicates for review. 

```{r}
# prepare data and recalculate duplicate counts
insps <-  insps |> 
  add_count(complaint_id,inspection_date,name="n_ids") |>
  add_count(complaint_id,inspection_date,inspection_type,name="n_ids_inspTypes") |>  
  mutate(diff_insp_types = n_ids!=n_ids_inspTypes,.after=inspection_type) |>  
  relocate(starts_with("x"),.after=last_col()) |>
  arrange(complaint_id,inspection_type)

lookup <- c(`Number of Duplicate IDs/Insp Dates`="n_ids",
            `All Insps Show Same Reinspection Status`="same_reinspection_status",
            `Number of Shared Violations`="n_common_viols",
            `Duplicates are Different Inspection Types?`="diff_insp_types")
insps |>
  filter(n_ids>1 & n_ids_inspTypes>1) |>
  select(-original_id,-root_id,-id,-inspection_result,-n_ids_inspTypes) |>
  relocate(inspection_date,.after=complaint_id) |>
  rename(all_of(lookup)) |>
  rename_with(\(x) str_to_title(str_replace_all(x,"_"," "))) |>
  make_logicals_easier() |>
  write_xlsx(paste(outputs_dir,
                   "Inspections on the Same Complaint and Day.xlsx",sep="/"))
```



## Export 
As noted above, not doing a full export, just exporting this version with basic checks, 
some deduplication, and cleaning.

```{r}
saveRDS(insps,paste(data_dir,"inspections_cleaner_dk.RDS",sep="/"))
```

